---
title: 'Filtering Data'
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Housekeeping
```{r housekeeping}
pacman::p_load(docstring, purrr, stringi, stringr, lubridate, geosphere, dplyr, data.table, dtplyr)
CODENAMES = c('Saitama', 'Sadaharu', 'Aizen', 'El Psy Kongroo')
```

## Learning Objective
Our goal is to filter observations by temporal, spatial, and topic dimensions.

## Scenario
Imagine that you are a spy and are trying to catch an enemy spy. Your intelligence network has compilated some strange transmissions for the past year and recorded the origin.

## Objective
Narrow down the possible leads to the enemy by filtering out the messages related to operational codewords within 5 kilometers of the suspected area taking into account when the enemy spy has been in town.

## The data

```{r helpers_data}
random_dt = function(n = 1e2,
                     st = as.Date(x = '2017-01-01'),
                     et = as.Date(x = '2017-12-31')) {
  #' Generates a series of n datetimes in the range of st and et
  #' 
  #' @param n how many datetimes to generate (default = 1e2)
  #' @param st start datetime (default = as.Date(x = '2017-01-01'))
  #' @param et end datetime (default = as.Date(x = '2017-12-31'))
  st = as.POSIXct(x = st)
  et = as.POSIXct(x = et)
  dt = as.numeric(x = difftime(time1 = et, time2 = st, units = 'sec'))
  ev = sort(x = runif(n = n, min = 0, max = dt))
  rt = st + ev
  }
secret_messages = function(p, codenames = CODENAMES) {
  #' Generates a message with p probability of including a codename
  #' 
  #' @param p probability of including a codename
  #' @param codenames possible codenames to include one
  if ((p <= 0) | (p >= 1)) {
    stop('p must a positive values less than one')
  }
  if (p > runif(n = 1)) {
    codename = sample(x = codenames, size = 1)
    secret = str_split(string = codename, pattern = ' ') %>%
      getElement(name = 1L) %>%
      str_c(collapse = stri_rand_strings(n = 1, length = 1))
    output = str_c(stri_rand_strings(n = 1,
                                     length = sample(x = 0:20,
                                                     size = 1)),
                   secret,
                   stri_rand_strings(n = 1,
                                     length = sample(x = 0:20,
                                                     size = 1)))
  } else {
      output = stri_rand_strings(n = 1,
                                 length = sample(20:35,
                                                 size = 1))
  }
  return(value = output)
  }
```

```{r data}
set.seed(seed = 0)
intelligence_report = data.table(
  datetime = random_dt(),
  longitude = runif(n = 100,
                    min = 38.880786 - 2,
                    max = 38.88078 + 2),
  latitude = runif(n = 100,
                   min = -77.115874 - 2e-1,
                   max = -77.115874 + 2e-1),
  message = map_chr(.x = runif(n = 100,
                               min = 1e-1,
                               max = 9e-1),
                    .f = secret_messages))
```

## Identify the following codenames in the messages

- Saitama
- Sadaharu
- Aizen
- El Psy Kongroo

If a codename has a space it may be masked by some random character so be aware of the masking techniques.

```{r messages}
intelligence_report = intelligence_report %>%
  mutate(codename = str_detect(string = message,
                               pattern = '(Saitama|Sadaharu|Aizen|El\\wPsy\\wKongroo)'))
```

## The suspected enemy HQ is located at 38.880786, -77.115874.

- Compute the distance between the origin of the message and the suspected enemy HQ

- Identify those that are nearby (in a 25 km area)

```{r spatial}
distances = apply(X = intelligence_report %>%
                            select(longitude, latitude),
                          MARGIN = 1,
                          FUN = function(lon_lat) {
                            distm(x = lon_lat,
                                  y = c(38.880786, -77.115874),
                                  fun = distVincentyEllipsoid) / 1e3
                            })
intelligence_report = intelligence_report %>%
  mutate(nearby = distances <= 25) # Is the origin nearby?
```

## Our reports indicate that the enemy spy likes to go the HQ Monday and Tuesday nights for TRIVIA night.

- Trivia night is Monday and Tuesday from 19:30 to 22:00

```{r trivia}
is_trivia = function(dt) {
  if (wday(x = dt) %in% 2:3) { # Is it Monday or Tuesday
    that_day = as.Date(x = dt)
    output = dt %within% interval(start = as.POSIXct(x = str_c(that_day,
                                                               '19:30')),
                                  end = as.POSIXct(x = str_c(that_day,
                                                             '22:00')))
  } else {
    output = FALSE
  }
  return(value = output)
  }
intelligence_report = intelligence_report %>%
  mutate(trivia = map_lgl(.x = datetime,
                          .f = is_trivia))
```

## Finding out the secret! Who is behind everything? Find the codename that matches the criteria

```{r decode}
output = intelligence_report %>%
  filter(codename & nearby & trivia) %>%
  getElement(name = 'message') %>%
  str_extract(pattern = '(Saitama|Sadaharu|Aizen|El\\wPsy\\wKongroo)')
```
